Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-12-12 paper.tex
plural

Note that the ambitions of our work stops at the level of
\emph{symmetric} cryptographic \emph{primitives} and, in particular,
their operational specification. These primitives would typically be
used (as black-boxes) in cryptographic \emph{protocols}: the analysis
of such protocols is out of the scope of the present work. Moreover,
we do not pretend that our conceptual apparatus could be useful to the
mathematical analysis of these primitives, as performed in a typical
``Section 2'', since this involves sophisticated, much more high-level
mathematical objects and results. Our focus is solely on shortening
the gap between an operational description of a primitive and its
efficient implementation. In effect, we are designing a programming
language dedicated to the description of symmetric cryptographic
primitives.

@@

For our present purposes, the essence of modern symmetric
cryptographic primitives can be distilled in a few paragraph. A
primitive can be understood as a purely functional program. It takes
two inputs: the ``plaintext'' and a list of ``round subkeys'' (which
are derived from a single key through a non-performance-critical
process called ``key schedule'' that we shall ignore here). The output
is the ``ciphertext''. The plaintext, the ciphertext and each
individual round subkeys are processed as ``blocks'' of binary data.
The block size is of the order of a hundred bits (typically, 64, 128
or 256 bits).

>>>>

Note that the ambitions of our work stop at the level of
\emph{symmetric} cryptographic \emph{primitives} and, in particular,
their operational specification. These primitives would typically be
used (as black-boxes) in cryptographic \emph{protocols}: the analysis
of such protocols is out of the scope of the present work. Moreover,
we do not pretend that our conceptual apparatus could be useful to the
mathematical analysis of these primitives, as performed in a typical
``Section 2'', since this involves sophisticated, much more high-level
mathematical objects and results. Our focus is solely on shortening
the gap between an operational description of a primitive and its
efficient implementation. In effect, we are designing a programming
language dedicated to the description of symmetric cryptographic
primitives.

@@

For our present purposes, the essence of modern symmetric
cryptographic primitives can be distilled in a few paragraphs. A
primitive can be understood as a purely functional program. It takes
two inputs: the ``plaintext'' and a list of ``round subkeys'' (which
are derived from a single key through a non-performance-critical
process called ``key schedule'' that we shall ignore here). The output
is the ``ciphertext''. The plaintext, the ciphertext and each
individual round subkeys are processed as ``blocks'' of binary data.
The block size is of the order of a hundred bits (typically, 64, 128
or 256 bits).
